Constructing hierarchical time series through clustering: \\Is there an optimal way for forecasting?

### 2024-09-03 Revision1/R1.tex
update

% \usepackage{natbib}
\usepackage{amsfonts,amsmath,bbm,bm,xcolor,booktabs,hyperref,amsthm}
\usepackage{geometry}
\usepackage{todonotes}
\usepackage{hyperref}
\usepackage{ulem}
\usepackage{subfig}
\geometry{a4paper,scale=0.8}
\usepackage{setspace}
\setstretch{1.5}
\let\code=\texttt
\let\proglang=\textsf
\setcounter{MaxMatrixCols}{20}
\usepackage{enumitem}
% \usepackage[pagewise]{lineno}

@@

@@

  \begin{keyword}
  Forecast reconciliation \sep
  Clustering \sep
  Hierarchy construction \sep
  Forecast combination \sep
  Mortality \sep
  Tourism flows
  \end{keyword}

@@

%The primary objective of hierarchical forecasting is to produce coherent forecasts for the series within a given HTS, which we refer to as \textit{original hierarchy}, ensuring that decision-makers at different levels of the hierarchy can analyze and make informed decisions based on aligned expectations for the future.
A limitation in the overwhelming majority of forecast reconciliation studies is that the structure of the hierarchy is taken as \textit{given}. This structure usually includes \textit{bottom level series}, an overall aggregate or \textit{top level series}, with various aggregation schemes used to construct \textit{middle level series}. Typically, middle levels are formed according to inherent attributes of the bottom-level series, such as geographical location, gender, product category, travel purpose, and others. We refer to this type of structure as the \textit{natural hierarchy}. While in some forecasting applications, decisions must be made with respect to the natural hierarchy, in other settings there might be some flexibility in determining how bottom levels are aggregated into middle levels. Moreover, a predefined natural hierarchy may not always exist in hierarchical settings, underscoring the importance of a hierarchical construction framework for forecasting purposes. It should be noted that very little attention has been paid to whether middle level series can be constructed in a way that leads to further improvements in forecast accuracy relative to a given natural hierarchy. To the best of our knowledge, only \cite{pangHierarchicalElectricityTime2018}, \cite{liForecastReconciliationApproach2019}, \cite{pangHierarchicalElectricityTime2022}, and \cite{matteraImprovingOutofSampleForecasts2023} have attempted to \textit{construct} middle level series in a data-driven way which ultimately improved forecast accuracy. However, all of these works use time series clustering to construct hierarchies in a manner that is somewhat ad hoc. Our work conducts a more thorough investigation of issues faced when  constructing hierarchical structures in forecast reconciliation. In particular, we address the following three research questions:%via time series clustering. %We also assess and disentangle the impact of

@@

\begin{researchquestion} In terms of forecast performance, can the use of middle level series lead to improvement compared to a two-level hierarchy (consisting of only top and bottom time series)? If so, is it possible to construct hierarchies in a data-driven way that leads to further improvements in forecast accuracy?

%In terms of forecast performance, is the natural hierarchy better than a two-level hierarchy (consisting of only top and bottom time series)? 
%In terms of forecast performance, does the natural hierarchy outperform a two-level hierarchy of only top and bottom time series?
\end{researchquestion}
%\begin{researchquestion}
%    If the use of middle level series in the ``natural'' hierarchy can lead to improvement in forecast accuracy, is it possible to construct hierarchies in a data-driven way that leads to further improvements in forecast accuracy?
%\end{researchquestion}

@@

The rationale behind the data-driven approach lies in grouping time series with similar patterns together, thereby creating middle-level series with enhanced signals and consequently, improved forecastability. Such arguments have been put forward by  \cite{pangHierarchicalElectricityTime2018}, \cite{liForecastReconciliationApproach2019}, \cite{pangHierarchicalElectricityTime2022}, and \cite{matteraImprovingOutofSampleForecasts2023}. They all demonstrate superior forecast performance from hierarchies constructed via clustering, relative to the natural hierarchy or the two-level hierarchy. However, these studies for the most part focus on a small number of (in some cases, a single) clustering techniques. In this paper, we take a more systematic approach by clustering time series using different representations (the original time series, forecast errors, features of both), different distance metrics (Euclidean, dynamic time warping), and different clustering paradigms ($k$-medioids, hierarchical). The models used to obtain base forecasts and the reconciliation method are fixed throughout the experiments. Using both empirical datasets, as well as a simulation study, we find evidence that constructing hierarchies via clustering can lead to improved forecasting performance, although the optimal clustering method depends on the dataset characteristics.% \textbf{We use the same base forecasting model and reconciliation method through this paper}.

While the idea behind time series clustering is intuitively appealing, the increased accuracy when using clustering-based methods may be attributed to two factors. The first, which we refer to as     ``grouping'' is the idea that some correct subsets of series are chosen to form new middle-level series. This is the argument commonly made to support clustering-based hierarchy construction \citep[see \textit{e.g.}][]{liForecastReconciliationApproach2019, pangHierarchicalElectricityTime2022, matteraImprovingOutofSampleForecasts2023}.  
%The idea that often motivates time series clustering is these ``groupings'' be composed of similar series. 
The second factor, which we refer to as the ``structure'' of the hierarchy, includes the number of middle-level series, the depth of the hierarchy, and the distribution of group sizes in the middle layer(s). Evidence showing that clustering within a forecast reconciliation framework leads to improved forecast accuracy, does not disentangle contributions from these two factors. Indeed, clustering methods may only work in so far as they generate a larger number of base forecasts. This argument would be consistent with the interpretation of reconciliation as a forecast combination of ``direct'' and ``indirect'' forecasts (\citealp{hollymanUnderstandingForecastReconciliation2021}), since more middle-level series implies a greater number of indirect forecasts in the combination. This leads to our second research question:

\begin{researchquestion}
    Should the improved accuracy of clustering-based methods be attributed to grouping together similar time series, or to the structure of the hierarchy? %including the number of middle level series?
\end{researchquestion}

@@

Note that forecast combination here differs from that of \cite{hollymanUnderstandingForecastReconciliation2021}, in that our approach averages not only different coherent forecasts, but also across hierarchies with completely different middle level series. This is possible since only coherent bottom and top level forecasts are averaged and evaluated.

@@

The rest of the paper is organized as follows. Section~\ref{sec:methodology} describes the trace minimization reconciliation methods employed and the clustering-based hierarchical time series augmentation techniques considered. Section~\ref{sec:cluster} first introduces the two datasets used, and then investigates RQ1, in particular the performance of cluster hierarchy compared to natural hierarchy and two-level hierarchy. Section~\ref{sec:permutation} introduces the novel permutation approach, followed by the investigation of RQ2 via evaluating the performance of natural hierarchy and the best performing cluster hierarchy compared to their respective random twins. To avoid the concern that clusters found in the empirical datasets are spurious, a simulation study is considered in Section~\ref{sec:simulation}. Section~\ref{sec:combination} covers the forecast combination approach raised in RQ3. Section~\ref{sec:conclusion} concludes this paper with discussions on the findings and outlines future research directions. 
% Data and code for reproducing the results in this paper are available at \url{https://github.com/AngelPone/project_hierarchy}.

@@

Hierarchical data can be characterized as being made up of bottom level series and their aggregates. In general, we consider a hierarchy with $n$ time series stacked into a vector $\boldsymbol{y}_t$. Let $\boldsymbol{b}_t$ denote the $m$ bottom level series, $\boldsymbol{a}_t$ denote the top level series, and $\boldsymbol{c}_t$ denote $k$ middle level series. The top level and bottom level will be common to all hierarchies we consider, and are augmented by middle level series. These middle level series can be formed according to attributes of the time series or in a data-driven fashion using time series clustering. The series are linked through an $(m+k+1)\times m$ summing matrix

%$\boldsymbol{S}$, which can be decomposed into an identity matrix $\boldsymbol{I}_m$ and two aggregation matrices $\boldsymbol{A}$ and $\boldsymbol{C}$ consisting of $0$ and $1$, such that, 
\[
  \boldsymbol{y}_t = \boldsymbol{S}\boldsymbol{b}_t = \begin{bmatrix}
    \boldsymbol{A} \\\boldsymbol{C} \\ \boldsymbol{I}_m 
  \end{bmatrix}  \boldsymbol{b}_t = \begin{bmatrix}
      \boldsymbol{a}_t \\ \boldsymbol{c}_t \\\boldsymbol{b}_t
  \end{bmatrix},
\]
where, $\boldsymbol{C}$ consists of zeros and ones that encode the aggregation, \textit{i.e.}, $c_{ij}=1$ if bottom level series $j$ is included in middle level series $i$, and zero otherwise. The top level series aggregates all bottom levels, \textit{i.e.}, $\boldsymbol{A} = \mathbf{1}_{1\times m}$, which is a row of ones.

@@

For the purposes of this paper, all forecasting is carried out as a two-step process. First, so-called \textit{base} forecasts are produced for all series in the hierarchy and stacked into an $n$-vector $\hat{\bm{y}}$, where subscripts are suppressed for brevity. For base forecasts, we use the Exponential Smoothing (ETS) method \citep{ForecastingExponentialSmoothing}, implemented using the \code{forecast} (\citealp{forecast}) package in {R} (\citealp{R}). Alternative methods such as ARIMA models were also considered, but this choice had very little impact on the overall conclusions.

The base forecasts generated in this first step, do not have the property that bottom level forecasts add up to forecasts of the aggregates, \textit{i.e.}, they are \textit{incoherent}. Therefore, forecast reconciliation is used as a post-forecasting step to ensure coherence of forecasts for all series in the hierarchy. In general, reconciliation takes the form of projecting the base forecasts as
\[
    \tilde{\boldsymbol{y}} = \boldsymbol{S}(\boldsymbol{S}'\boldsymbol{W}_h^{-1}\boldsymbol{S})^{-1}\boldsymbol{S}'\boldsymbol{W}_h^{-1}\hat{\boldsymbol{y}},
\]
where $\tilde{\boldsymbol{y}}$ are the \textit{reconciled} forecasts and $\boldsymbol{W}_h$ is the covariance matrix of $h$-step-ahead forecast errors. For reconciliation, we use the MinT method \citep{wickramasuriyaOptimalForecastReconciliation2019}, which involves plugging in a shrinkage estimator for $\boldsymbol{W}_h$. Although alternatives were considered, including approaches that assume $\boldsymbol{W}_h$ is diagonal or an identity matrix \citep{hyndmanOptimalCombinationForecasts2011}, this choice has little impact on our main conclusions.

@@

To the best of our knowledge, four studies have attempted to improve forecast accuracy in a reconciliation setting by constructing middle levels of the hierarchy using time series clustering.
\cite{pangHierarchicalElectricityTime2018} detect consumption patterns of electricity smart meter data based on X-means clustering algorithm, while \cite{pangHierarchicalElectricityTime2022} propose several alternative clustering methods to group similar electricity and solar power time series. 
\cite{liForecastReconciliationApproach2019} apply agglomerative hierarchical clustering to cause-of-death time series, and
\cite{matteraImprovingOutofSampleForecasts2023} utilize Partition Around Medoids algorithms to unveil underlying structures in stock price indexes. 
However, these studies are limited in scope as they focus on a few  clustering techniques. 
%\todo{Should we move to intro?}
%construct hierarchies through clustering algorithms and 
Inspired by the comprehensive overview of time series clustering by \cite{aghabozorgiTimeseriesClusteringDecade2015a}, we consider various approaches based on three key components, namely \textit{time series representations}, \textit{distance measures}, and \textit{clustering algorithms}. %Our framework for creating a clustering approach is as follows: we first choose a time series representation, which can be the raw time series, the in-sample forecast error, features of the time series, or features of the in-sample forecast error. We then select a distance measure, either Euclidean distance or dynamic time warping. Once the representation and the distance measure are determined, we choose a clustering algorithm, either $k$-medoids or agglomerative hierarchical clustering, to form clusters in the dataset. In total, we consider 12 different clustering approaches in our experiments.
%, and evaluation measures.

@@

The time series representation refers to the object that acts as an input for time series clustering. %that provide diverse perspectives on the same dataset.
% Considering the impracticality of exploring every conceivable time series representation, we focus on four key ones: raw time series without any transformation (hereafter referred to as ``time series''), in-sample one-step-ahead forecast error, features of time series, and features of in-sample one-step-ahead forecast error.
Our first candidate for the representation is raw time series itself, due to its simplicity and broad applicability.
%However, a crucial element contributing to the success of the MinT method lies in estimating the covariance matrix of base forecast errors based on in-sample one-step-ahead forecast error. 
We also consider the in-sample one-step-ahead forecast error as a representation of the time series, since a key step in MinT reconciliation is to estimate the $\boldsymbol{W}_h$ matrix. %to examine how the structure within the error series influences reconciliation performance. 
It is important to note that raw time series and in-sample error representations are standardized to eliminate the impact of scale variations. %This is done by normalizing each series by subtracting its mean and then dividing by the standard error.

A potential shortcoming to using the time series or forecast error as a representation is their high dimensionality, which is equal to the sample size of the training data. To address this concern, low dimension summaries of ``features'' can be considered. Features have been used in the context
%Features, widely employed in capturing time series characteristics, play a pivotal role in various time series applications, including 
time series clustering by \cite{tianoFeatTSFeaturebasedTime2021}, and for forecasting by \cite{wangUncertaintyEstimationFeaturebased2022} and \cite{ liFeaturebasedIntermittentDemand2023}. 
We consider features of both the raw time series and the in-sample forecast error as representations. After filtering out the features that are constant across all series, we select $56$ features calculated by the \code{tsfeatures} (\citealp{tsfeatures}) package in R. %have been applied in several feature-based forecasting studies, \textit{e.g.}, \cite{montero-mansoFFORMAFeaturebasedForecast2020}.
To the best of our knowledge, we are the first to utilize in-sample forecast error and time series features as representations in the context of forecast reconciliation.
%\sout{These representations allow us to gain insights into the diverse aspects of hierarchical time series data}.

@@

\paragraph{\textbf{Clustering algorithms}}
%\sout{Clustering algorithms, such as partitioning, hierarchical, grid-based, and model-based approaches,  identify clusters by optimizing specific objective functions} \citep{aghabozorgiTimeseriesClusteringDecade2015a}. %, or applying heuristics. %These functions are calculated based on chosen time series representations and distance measures.
In this paper, we focus on two clustering algorithms, namely $k$-medoids and agglomerative hierarchical clustering. These algorithms are implemented using the \texttt{cluster} (\citealp{cluster}) package in R. {The $k$-medoids algorithm aims to minimize the total distance between all observations within a cluster and their respective cluster median.} This is implemented using the partitioning around medoids (PAM) method (\citealp{PartitioningMedoidsProgram1990}).
%Unlike $k$-means which employs the mean vector of samples as the cluster center, $k$-medoids selects one observation within each cluster as the center. 
Following the recommendation of \cite{PartitioningMedoidsProgram1990}, we determine the optimal number of clusters using the average silhouette width (ASW), which is a commonly used index in cluster validation \citep[see \textit{e.g.},][]{shutaywi2021silhouette}.
%ASW assesses the quality of clustering results by measuring the proximity of samples within a cluster compared to neighboring clusters.
%Given a clustering result, the silhouette width for the $i$th sample is calculated as 
%\[
%  SW(i) = \frac{b(i)-a(i)}{\max\{a(i), b(i)\}},  
%\]
%where $a(i)$ represents the average distance of the $i$th observation to others in the same cluster, and $b(i)$ is the average distance to observations in the cluster nearest to observation excluding the cluster that observation $i$ is assigned to. A higher silhouette width indicates greater proximity to observations within the same cluster than to those in neighboring clusters. The ASW is the average of silhouette width across all observation. We select the clustering result that maximizes the ASW by iterating over all possible numbers of clusters. However, ASW has the limitation of being undefined when there is only one cluster. %\textbf{Gap statistics (\citealp{tibshiraniEstimatingNumberClusters2002}) can be used to address this limitation, but it is incompatible with DTW.}

Agglomerative hierarchical clustering initializes each observation in its own cluster, and then merges the nearest two clusters in a stepwise fashion until all observations form a single cluster. We employ Ward's linkage (\citealp{murtaghWardHierarchicalAgglomerative2014a}) which defines the nearest clusters as those that minimize the increase in within-cluster variance at each step. Applying hierarchical clustering to $m$ bottom-level series results in a binary hierarchical tree with ($2m-1$) nodes, all of which are retained as middle level series.

\begin{figure}[h!]
    \centering
    \vspace{0.2in}
    \includegraphics[width=0.46\textwidth]{figures/pamcluster.pdf}
    \hspace{1cm}
    \includegraphics[width=0.46\textwidth]{figures/aggcluster.pdf}
    \caption{\label{fig:cluster_example}Example clustering results of two clustering algorithms. Left panel displays example for $k$-medoids algorithm, and right panel displays example for agglomerative hierarchical clustering algorithm.}
\end{figure}
Figure~\ref{fig:cluster_example} illustrates two examples of hierarchies generated by $k$-medoids and agglomerative clustering algorithms, on the left and right panels respectively. The corresponding two-level hierarchy would have four bottom-level series and one top-level series. % \sout{These examples highlight the distinct behaviors of the two algorithms.}
Note that, $k$-medoids constructs a simple hierarchy with a single middle level, while hierarchical clustering generates multiple nested middle levels. 
In general, $k$-medoids produces a hierarchy with fewer middle levels and middle level series compared to hierarchical clustering. As the number of bottom-level series increases, these differences become increasingly pronounced, with potential implications for forecast reconciliation.

>>>>

% \usepackage{natbib}
\usepackage{amsfonts,amsmath,bbm,bm,xcolor,booktabs,hyperref,amsthm}
\usepackage{geometry}
\usepackage{todonotes}
\usepackage{hyperref}
\usepackage{ulem}
\usepackage{subfig}
\geometry{a4paper,scale=0.8}
\usepackage{setspace}
\setstretch{1.5}
\let\code=\texttt
\let\proglang=\textsf
\setcounter{MaxMatrixCols}{20}
\usepackage{enumitem}

@@

    %the dynamics between ``grouping'' and ``structure'', which lead to an improved understanding of forecast reconciliation.\\

@@

  \begin{keyword}
  Forecast reconciliation \sep
  Hierarchical time series \sep
  Clustering \sep
  Hierarchy construction \sep
  Forecast combination
  \end{keyword}

@@

%The primary objective of hierarchical forecasting is to produce coherent forecasts for the series within a given HTS, which we refer to as \textit{original hierarchy}, ensuring that decision-makers at different levels of the hierarchy can analyze and make informed decisions based on aligned expectations for the future.
A limitation in the overwhelming majority of forecast reconciliation studies is that the structure of the hierarchy is taken as \textit{given}. This structure usually includes \textit{bottom level series}, an overall aggregate or \textit{top level series}, with various aggregation schemes used to construct \textit{middle-level series}. Typically, middle levels are formed according to inherent attributes of the bottom-level series, such as geographical location, gender, product category, travel purpose, and others. We refer to this type of structure as the \textit{natural hierarchy}. While in some forecasting applications, decisions must be made with respect to the natural hierarchy, in other settings there might be some flexibility in determining how bottom levels are aggregated into middle levels. Moreover, a predefined natural hierarchy may not always exist in hierarchical settings, underscoring the importance of a hierarchical construction framework for forecasting purposes. It should be noted that very little attention has been paid to whether middle-level series can be constructed in a way that leads to further improvements in forecast accuracy relative to a given natural hierarchy. {\color{purple}To the best of our knowledge, only \cite{yang2017reconciling}, \cite{pangHierarchicalElectricityTime2018}, \cite{liForecastReconciliationApproach2019}, \cite{pangHierarchicalElectricityTime2022}, \cite{cini2023graph}, and \cite{matteraImprovingOutofSampleForecasts2023} have attempted to \textit{construct} middle-level series in a data-driven way which ultimately improved forecast accuracy. } However, all of these works use time series clustering to construct hierarchies in a manner that is somewhat ad hoc. Our work conducts a more thorough investigation of issues faced when  constructing hierarchical structures in forecast reconciliation. In particular, we address the following three research questions:%via time series clustering. %We also assess and disentangle the impact of

@@

\begin{researchquestion} In terms of forecast performance, can the use of middle-level series lead to improvement compared to a two-level hierarchy (consisting of only top and bottom time series)? If so, is it possible to construct hierarchies in a data-driven way that leads to further improvements in forecast accuracy?

%In terms of forecast performance, is the natural hierarchy better than a two-level hierarchy (consisting of only top and bottom time series)? 
%In terms of forecast performance, does the natural hierarchy outperform a two-level hierarchy of only top and bottom time series?
\end{researchquestion}
%\begin{researchquestion}
%    If the use of middle-level series in the ``natural'' hierarchy can lead to improvement in forecast accuracy, is it possible to construct hierarchies in a data-driven way that leads to further improvements in forecast accuracy?
%\end{researchquestion}

@@

The rationale behind the data-driven approach lies in grouping time series with similar patterns together, thereby creating middle-level series with enhanced signals and consequently, improved forecastability. Such arguments have been put forward by  \cite{pangHierarchicalElectricityTime2018}, \cite{liForecastReconciliationApproach2019}, \cite{pangHierarchicalElectricityTime2022}, and \cite{matteraImprovingOutofSampleForecasts2023}. They all demonstrate superior forecast performance from hierarchies constructed via clustering, relative to the natural hierarchy or the two-level hierarchy. However, these studies for the most part focus on a small number of (in some cases, a single) clustering techniques. In this paper, we take a more systematic approach by clustering time series using different representations (the original time series, forecast errors, features of both), different distance metrics (Euclidean, dynamic time warping), and different clustering paradigms ($k$-medioids, hierarchical). The models used to obtain base forecasts and the reconciliation method are fixed throughout the experiments. Using both empirical datasets, as well as a simulation study, we find evidence that constructing hierarchies via clustering can lead to improved forecasting performance, although the optimal clustering method depends on the dataset as well as the base forecasting and reconciliation method.% \textbf{We use the same base forecasting model and reconciliation method through this paper}.

While the idea behind time series clustering is intuitively appealing, the increased accuracy when using clustering-based methods may be attributed to two factors. The first, which we refer to as     ``grouping'' is the idea that some correct subsets of series are chosen to form new middle-level series. This is the argument commonly made to support clustering-based hierarchy construction \citep[see \textit{e.g.}][]{liForecastReconciliationApproach2019, pangHierarchicalElectricityTime2022, matteraImprovingOutofSampleForecasts2023}.  
%The idea that often motivates time series clustering is these ``groupings'' be composed of similar series. 
The second factor, which we refer to as the ``structure'' of the hierarchy, includes the number of middle-level series, the depth of the hierarchy, and the distribution of group sizes in the middle layer(s). Evidence showing that clustering within a forecast reconciliation framework leads to improved forecast accuracy, does not disentangle contributions from these two factors. Indeed, clustering methods may only work in so far as they generate a larger number of base forecasts. {\color{purple}This argument would be consistent with the interpretation of reconciliation as a forecast combination of ``direct'' and ``indirect'' forecasts (\citealp{hollymanUnderstandingForecastReconciliation2021, di2024forecast}), since more middle-level series implies a greater number of indirect forecasts in the combination.} This leads to our second research question:

\begin{researchquestion}
    Should the improved accuracy of clustering-based methods be attributed to grouping together similar time series, or to the structure of the hierarchy? %including the number of middle-level series?
\end{researchquestion}

@@

Note that forecast combination here differs from that of \cite{hollymanUnderstandingForecastReconciliation2021}, in that our approach averages not only different coherent forecasts, but also across hierarchies with completely different middle-level series. This is possible since only coherent bottom and top level forecasts are averaged and evaluated.

@@

The rest of the paper is organized as follows. Section~\ref{sec:methodology} describes the trace minimization reconciliation methods employed and the clustering-based hierarchical time series augmentation techniques considered. Section~\ref{sec:cluster} first introduces the two datasets used, and then investigates RQ1, in particular the performance of cluster hierarchy compared to natural hierarchy and two-level hierarchy. Section~\ref{sec:permutation} introduces the novel permutation approach, followed by the investigation of RQ2 via evaluating the performance of natural hierarchy and the best performing cluster hierarchy compared to their respective random twins. To avoid the concern that clusters found in the empirical datasets are spurious, a simulation study is considered in Section~\ref{sec:simulation}. Section~\ref{sec:combination} covers the forecast combination approach raised in RQ3. Section~\ref{sec:conclusion} concludes this paper with discussions on the findings and outlines future research directions. 
Data and code for reproducing the results in this paper are available at \url{https://github.com/Purity3-cyber/constructing_hierarchies}.

@@

Hierarchical data can be characterised as being made up of bottom level series and their aggregates. In general, we consider a hierarchy with $n$ time series stacked into a vector $\boldsymbol{y}_t$. Let $\boldsymbol{b}_t$ denote the $m$ bottom level series, $\boldsymbol{a}_t$ denote the top level series, and $\boldsymbol{c}_t$ denote $k$ middle-level series. The top level and bottom level will be common to all hierarchies we consider, and are augmented by middle-level series. These middle-level series can be formed according to attributes of the time series or in a data-driven fashion using time series clustering. The series are linked through an $(m+k+1)\times m$ summing matrix 
%$\boldsymbol{S}$, which can be decomposed into an identity matrix $\boldsymbol{I}_m$ and two aggregation matrices $\boldsymbol{A}$ and $\boldsymbol{C}$ consisting of $0$ and $1$, such that, 
\[
  \boldsymbol{y}_t = \boldsymbol{S}\boldsymbol{b}_t = \begin{bmatrix}
    \boldsymbol{A} \\\boldsymbol{C} \\ \boldsymbol{I}_m 
  \end{bmatrix}  \boldsymbol{b}_t = \begin{bmatrix}
      {\color{purple}a_t} \\ \boldsymbol{c}_t \\\boldsymbol{b}_t
  \end{bmatrix},
\]
where, $\boldsymbol{C}$ consists of zeros and ones that encode the aggregation, \textit{i.e.}, $c_{ij}=1$ if bottom level series $j$ is included in middle-level series $i$, and zero otherwise. The top level series aggregates all bottom levels, \textit{i.e.}, $\boldsymbol{A} = \mathbf{1}_{1\times m}$, which is a row of ones.

@@

For the purposes of this paper, all forecasting is carried out as a two-step process. First, so called \textit{base} forecasts are produced for all series in the hierarchy and stacked into an $n$-vector $\hat{\bm{y}}$, where subscripts are suppressed for brevity. For base forecasts, we use the Exponential Smoothing (ETS) method \citep{ForecastingExponentialSmoothing}, implemented using the \code{forecast} (\citealp{forecast}) package in {R} (\citealp{R}). Alternative methods such as ARIMA models were also considered, but this choice had very little impact on the overall conclusions.

The base forecasts generated in this first step, do not have the property that bottom level forecasts add up to forecasts of the aggregates , \textit{i.e.}, they are \textit{incoherent}. Therefore, forecast reconciliation is used as a post-forecasting step to ensure coherence of forecasts for all series in the hierarchy. In general, reconciliation takes the form of projecting the base forecasts as
\vspace{-0.1in}
\[
  \color{purple}  \tilde{\boldsymbol{y}} = \boldsymbol{S}(\boldsymbol{S}'\boldsymbol{W}^{-1}\boldsymbol{S})^{-1}\boldsymbol{S}'\boldsymbol{W}^{-1}\hat{\boldsymbol{y}},
  \vspace{-0.15in}
\]
where $\tilde{\boldsymbol{y}}$ are the \textit{reconciled} forecasts and $\boldsymbol{W}$ is the covariance matrix of base forecast errors. {\color{purple}For reconciliation, we use the MinT method \citep{wickramasuriyaOptimalForecastReconciliation2019} implemented using the \texttt{FoReco} (\citealp{FoReco}) package in R, which involves plugging in a shrinkage estimator for $\boldsymbol{W}$.}Although alternatives were considered, including approaches that assume $\boldsymbol{W}$ is diagonal or an identity matrix \citep{hyndmanOptimalCombinationForecasts2011}, this choice has little impact on our main conclusions. {\color{purple}Despite the importance of non-negativity in practical scenarios such as demand forecasting, our study does not enforce non-negative forecasts for two reasons. First, the proportion of negative forecasts is minimal in our applications (0.25\% in the tourism dataset and 0.2\% in the mortality dataset). Second, introducing non-negativity constraints would significantly increase the computational cost of our experiments. It should be noted that the framework proposed in our paper could be applied using any reconciliation approach, including those that guarantee non-negative forecasts.}

@@

%To the best of our knowledge,

%four studies

{\color{purple}Only a handful of studies have attempted to improve forecast accuracy in a reconciliation setting by constructing middle levels of the hierarchy using time series clustering. \cite{yang2017reconciling} employ spatial $k$-means clustering methods to assign distributed photovoltaics  to their nearest transmission grid bus locations for reconciliation purposes.} \cite{pangHierarchicalElectricityTime2018} detect consumption patterns of electricity smart meter data based on X-means clustering algorithm, while \cite{pangHierarchicalElectricityTime2022} propose several alternative clustering methods to group similar electricity and solar power time series. 
\cite{liForecastReconciliationApproach2019} apply agglomerative hierarchical clustering to cause-of-death time series, and
\cite{matteraImprovingOutofSampleForecasts2023} utilize Partition Around Medoids algorithms to unveil underlying structures in stock price indexes. {\color{purple} \cite{cini2023graph} took an end-to-end approach from which clusters can be extracted together with hierarchical forecasts.} 
However, these studies are limited in scope as they focus on a small number of  clustering techniques. 
%\todo{Should we move to intro?}
%construct hierarchies through clustering algorithms and 
Inspired by the comprehensive overview of time series clustering by \cite{aghabozorgiTimeseriesClusteringDecade2015a}, we consider various approaches based on three key components, namely \textit{time series representations}, \textit{distance measures}, and \textit{clustering algorithms}. %Our framework for creating a clustering approach is as follows: we first choose a time series representation, which can be the raw time series, the in-sample forecast error, features of the time series, or features of the in-sample forecast error. We then select a distance measure, either Euclidean distance or dynamic time warping. Once the representation and the distance measure are determined, we choose a clustering algorithm, either $k$-medoids or agglomerative hierarchical clustering, to form clusters in the dataset. In total, we consider 12 different clustering approaches in our experiments.
%, and evaluation measures.

@@

The time series representation refers to the object that acts as an input for time series clustering. %that provide diverse perspectives on the same dataset.
% Considering the impracticality of exploring every conceivable time series representation, we focus on four key ones: raw time series without any transformation (hereafter referred to as ``time series''), in-sample one-step-ahead forecast error, features of time series, and features of in-sample one-step-ahead forecast error.
{\color{purple} Our first candidate for the representation, due to its simplicity and broad applicability, is the time series itself without any processing or additional features. We refer to this as the raw time series.}
%However, a crucial element contributing to the success of the MinT method lies in estimating the covariance matrix of base forecast errors based on in-sample one-step-ahead forecast error. 
We also consider the in-sample one-step-ahead forecast error as a representation of the time series, since a key step in MinT reconciliation is to estimate the $\boldsymbol{W}$ matrix. %to examine how the structure within the error series influences reconciliation performance. 
It is important to note that raw time series and in-sample error representations are standardized to eliminate the impact of scale variations. %This is done by normalizing each series by subtracting its mean and then dividing by the standard error.

A potential shortcoming to using the time series or forecast error as a representation is their high dimensionality, which is equal to the sample size of the training data. To address this concern, low dimension summaries of ``features'' can be considered. Features have been used in the context {\color{purple} of}
%Features, widely employed in capturing time series characteristics, play a pivotal role in various time series applications, including 
time series clustering by \cite{tianoFeatTSFeaturebasedTime2021}, and for forecasting by \cite{wangUncertaintyEstimationFeaturebased2022} and \cite{ liFeaturebasedIntermittentDemand2023}. 
We consider features of both the raw time series and the in-sample forecast error as representations. After filtering out the features that are constant across all series, we select $56$ features.\footnote{The list and descriptions of features are available in the online GitHub repository.} These time series features are calculated by the \code{tsfeatures} (\citealp{tsfeatures}) package in R. %have been applied in several feature-based forecasting studies, \textit{e.g.}, \cite{montero-mansoFFORMAFeaturebasedForecast2020}.
{\color{purple}To the best of our knowledge, we are the first to utilize in-sample forecast error and time series features as representations in the clustering-based hierarchy construction process.\footnote{We acknowledge that in-sample forecast error and time series features have been used in the computation of base forecasts or in the reconciliation process \citep[see][]{wickramasuriyaOptimalForecastReconciliation2019}.}}
%To the best of our knowledge, we are the first to utilize in-sample forecast error and time series features as representations in the context of forecast reconciliation.
%\sout{These representations allow us to gain insights into the diverse aspects of hierarchical time series data}.

@@

\paragraph{\textbf{Clustering algorithms}}
%\sout{Clustering algorithms, such as partitioning, hierarchical, grid-based, and model-based approaches,  identify clusters by optimizing specific objective functions} \citep{aghabozorgiTimeseriesClusteringDecade2015a}. %, or applying heuristics. %These functions are calculated based on chosen time series representations and distance measures.
In this paper, we focus on two clustering algorithms, namely $k$-medoids and agglomerative hierarchical clustering. These algorithms are implemented using the \texttt{cluster} (\citealp{cluster}) package in R. {The $k$-medoids algorithm aims to minimize the total distance between all observations within a cluster and their respective cluster median.} This is implemented using the partitioning around medoids (PAM) method (\citealp{PartitioningMedoidsProgram1990}).
%Unlike $k$-means which employs the mean vector of samples as the cluster center, $k$-medoids selects one observation within each cluster as the center. 
Following the recommendation of \cite{PartitioningMedoidsProgram1990}, we determine the optimal number of clusters using the average silhouette width (ASW), which is a commonly used index in cluster validation \citep[see \textit{e.g.}, ][]{shutaywi2021silhouette}.
%ASW assesses the quality of clustering results by measuring the proximity of samples within a cluster compared to neighboring clusters.
%Given a clustering result, the silhouette width for the $i$th sample is calculated as 
%\[
%  SW(i) = \frac{b(i)-a(i)}{\max\{a(i), b(i)\}},  
%\]
%where $a(i)$ represents the average distance of the $i$th observation to others in the same cluster, and $b(i)$ is the average distance to observations in the cluster nearest to observation excluding the cluster that observation $i$ is assigned to. A higher silhouette width indicates greater proximity to observations within the same cluster than to those in neighboring clusters. The ASW is the average of silhouette width across all observation. We select the clustering result that maximizes the ASW by iterating over all possible numbers of clusters. However, ASW has the limitation of being undefined when there is only one cluster. %\textbf{Gap statistics (\citealp{tibshiraniEstimatingNumberClusters2002}) can be used to address this limitation, but it is incompatible with DTW.}

Agglomerative hierarchical clustering initialises each observation in its own cluster, and then merges the nearest two clusters in a stepwise fashion until all observations form a single cluster. We employ Ward's linkage (\citealp{murtaghWardHierarchicalAgglomerative2014a}) which defines the nearest clusters as those that minimize the increase in within-cluster variance at each step. Applying hierarchical clustering to $m$ bottom-level series results in a binary hierarchical tree with ($2m-1$) nodes, all of which are retained as middle-level series.

\begin{figure}[h!]
    \centering
  %  \vspace{-0.1in}
    \includegraphics[width=0.46\textwidth]{../manuscript/figures/pamcluster.pdf}
    \hspace{1cm}
    \includegraphics[width=0.46\textwidth]{../manuscript/figures/aggcluster.pdf}
    \caption{\label{fig:cluster_example}Example clustering results of two clustering algorithms. Left panel displays example for $k$-medoids algorithm, and right panel displays example for agglomerative hierarchical clustering algorithm.}
\end{figure}
